### 2008-03-19 trunk/Papers/dl-gre-08/gre.tex
simplified the picture
Written the algorithm for ALC and changed the one for EL.

\begin{algorithm}[t]
\dontprintsemicolon
\caption{\el\ bisimulation classes}
\label{algo:bisim-el}
$\C \leftarrow \{\top\}$ \;
\For{$p \in \prop$}{
  add$_\el$(p, \C) \;
}
\While{not all concepts in $\C$ singleton}{
  \For{$C \in \C$, $R \in \rel$}{
    add$_\el$($\exists R.C$, $\C$) \;
  }
  \If{made no changes to $\C$}{
    exit\;
  }
}
\end{algorithm}

\begin{algorithm}[t]
\dontprintsemicolon
\caption{add$_\el$($C$, $\C$)}
\label{algo:bisim-add-el}
\For{$C' \in \C$ non-singleton}{
  \If{$C' \sqcap C$ is not subsumed in $\C$}{
    add $C' \sqcap C$ to $\C$ \;
    remove subsumed concepts from $\C$\;
  }
}
\end{algorithm}

@@

Compared to the \alc\ algorithm, the \el\ algorithm will generally
compute larger bisimulation classes, and compute singleton
bisimulation classes less frequently, because the concepts it can
build can ``see'' fewer differences between individuals.  For example,
in the model shown in Fig.~\ref{fig:el-vs-alc}, the \alc\ algorithm
will compute the concepts $\{\neg \exists R. (\neg P \sqcap \neg Q),
\exists R. (\neg P \sqcap \neg Q), Q, \neg P \sqcap \neg Q\}$ (each of
which is singleton), but $a$ and $b$ are \el-bisimilar in this model,
and the only \el-concept that is true in $d$ is $\top$, so the \el\
algorithm can only compute the concepts $\{\top, P, P \sqcap Q\}$.
Furthermore, because the \el\ algorithm maintains a weaker invariant
for the classes, it may consider more of these classes simultaneously.
Prima facie, this means that it has worst-case exponential runtime,
given that the whole domain has an exponential number of subsets.
However, it is possible that a more careful analysis of the
combinatorics of the considered subsets reveal that the algorithm is
actually polynomial; it certainly behaves polynomially in our
evaluations (see below).

>>>>

\begin{algorithm}[t]
\dontprintsemicolon
\caption{\el\ bisimulation classes}
\label{algo:bisim-el}
\KwIn{A model $\gM = (\Delta^\gM, |\cdot|^\gM)$}
\KwOut{A set \RE of pairs $(\varphi,|\varphi|^\gM)$ such that
$\{S \mid (\varphi,S) \in \RE\}$ is the set of $\el$-bisimulation
classes of $\gM$.}
\RE $\leftarrow \{(\top,\Delta^\gM)\}$ \;
\For{$p \in \prop$}{
  add$_\el$(p, \RE) \;
}
\While{for some $(\varphi,S) \in \RE$, $|S| > 1$}{
  \For{$(\varphi,S) \in$ \RE, $R \in \rel$}{
    $\RE \leftarrow$ add$_\el$($\exists R.\varphi$, \RE) \;
  }
  \If{made no changes to $\RE$}{
    exit\;
  }
}
\end{algorithm}

\begin{algorithm}[t]
\dontprintsemicolon
\caption{add$_\el$($\varphi$, $\RE$)}
\label{algo:bisim-add-el}
\For{$(\psi,S) \in \RE$ with $|S| > 1$}{
  \If{$\psi \sqcap \varphi$ is not subsumed in $\RE$ {\bf and}
    $|\psi \sqcap \varphi|^\gM \neq \emptyset$ {\bf and}
    $|\psi \sqcap \varphi|^\gM \neq |\psi|^\gM$}{
    add $(\psi \sqcap \varphi,|\psi \sqcap \varphi|^\gM)$ to $\RE$ \;
    remove subsumed concepts from $\RE$\;
  }
}
\end{algorithm}

\begin{algorithm}[t]
\dontprintsemicolon
\caption{}
\label{algo:bisim-alc}
\KwIn{A model $\gM = (\Delta^\gM, |\cdot|^\gM)$}
\KwOut{A set \RE of pairs $(\varphi,|\varphi|^\gM)$ such that
$\{S \mid (\varphi,S) \in \RE\}$ is the set of $\alc$-bisimulation
classes of $\gM$.}
$\RE \leftarrow \{(\top,\Delta^\gM)\}$

\For{$p \in \prop$}{
   \For{$(\varphi,S) \in \RE$}{
      split$((\varphi,S),p,\RE)$
   }
}

\While{for some $(\varphi,S) \in \RE, |S|>1$}{
   \For{$(\varphi,S) \in \RE, R \in \rel$}{
      \For{any set $\{(\psi_1,S_1),\ldots,(\psi_n,S_n)\} \in \RE$}{
         split$((\varphi,S),\exists R.\bigsqcup_i \psi_i,\RE)$
         }
   }
   \If{made no changes to \RE}{
      exit
      }
}

\end{algorithm}

\begin{algorithm}
\caption{split$((\varphi,S),\psi, \RE)$}
   \If{$S \cap |\psi|^\gM \not = \emptyset$ and
    $S \backslash |\psi|^\gM \not = \emptyset$}{
         add  $(\varphi \sqcap \psi     , S \cap |\psi|^\gM)$ and
               $(\varphi \sqcap \neg \psi, S \backslash |\psi|^\gM)$ to \RE \;
         delete $(\varphi,S)$ from \RE \;
      }
\end{algorithm}

@@

Compared to the \alc\ algorithm, the \el\ algorithm will generally
compute larger bisimulation classes, and compute singleton
bisimulation classes less frequently, because the concepts it can
build can ``see'' fewer differences between individuals.  For example,
in the model shown in Fig.~\ref{fig:el-vs-alc}, the \alc\ algorithm
will compute the concepts $\{\neg \exists R. (\neg P \sqcap \neg Q),
Q, \exists R. (\neg P \sqcap \neg Q), \neg P \sqcap \neg Q\}$ (each of
which is singleton), but $a$ and $c$ are \el-bisimilar in this model,
and the only \el-concept that is true in $d$ is $\top$, so the \el\
algorithm can only compute the concepts $\{\top, P, P \sqcap Q\}$.
Furthermore, because the \el\ algorithm maintains a weaker invariant
for the classes, it may consider more of these classes simultaneously.
Prima facie, this means that it has worst-case exponential runtime,
given that the whole domain has an exponential number of subsets.
However, it is possible that a more careful analysis of the
combinatorics of the considered subsets reveal that the algorithm is
actually polynomial; it certainly behaves polynomially in our
evaluations (see below).
